---
title: "In-class Exercise 11-3"
subtitle: "North Carolina Voting"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

## Polling Questions

In a poll conducted by Reuters / IPSOS, they take a random sample of 1854 likely 
voters and fidn that 52.1% support Biden and 47.8% support Trump.

Is there enough evidence to claim that a majority of likely voters support
Biden based on this sample?

1. Define the hypotheses.

$Ho$: The true proportion of likely voters ewho are planning to vote for Biden 
is equal to 0.50.

$Ha$: The true proportion of likely voters who are planning to vote for Biden is 
greater than 0.50.

$Ho: p \leq 0.50$

$Ha: p > 0.50$

2. Collect and summarize data.

```{r sample-phat}
phat <- 563 / 1080
phat
```

3. Assess the evidence.

Find the probability of getting sample data as or more extreme as ours, assuming 
the null hypothesis is true (assuming $Ho: p = 0.50$). This is the $p$-value.

```{r sim-p-value}

set.seed(1)
num_sims <- 10000

sim_dist <- numeric(num_sims)
outcomes <- c("Biden", "Trump")

for(i in 1:num_sims){
  sim <- sample(outcomes, size = 1080, prob = c(0.50, 0.50), replace = TRUE)
  sim_dist[i] <- sum(sim == "Biden") / length(sim)
}

null_dist <- tibble(sim_dist)

ggplot(null_dist,
       aes(x = sim_dist)) +
  geom_histogram() +
  geom_vline(xintercept = phat, lwd = 2, col = "red")

null_dist %>%
  summarize(p_value = mean(sim_dist >= phat))

```


```{r packages, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(sf)
```

We first read in the file `nc_votes.shp` with the `st_read` function, saving as
`nc_votes`.

```{r}
nc_votes <- st_read("data/nc_votes.shp", quiet = TRUE)
```

1. Carefully examine `nc_votes`. How does it differ from the data frames we have
seen before? What type of geometry is associated with each simple feature?

```{r ex-1}
nc_votes
```


2. Use `nc_votes` to create a map of North Carolina.

```{r ex-2}
ggplot(nc_votes) + 
  geom_sf()
```


3. Build on your previous map so that the counties have a fill color 
corresponding to the percentage of mail-in ballots that were rejected. Carefully
consider effective labels, colorings, and aesthetics.

```{r ex-3}
ggplot(nc_votes) + 
  geom_sf(aes(fill = rejectd / mailed)) +
  scale_fill_gradient(low = "#fde0dd", high = "#c51b8a") +
  theme_bw() +
  labs(fill = "Proportion of mailed ballots that were rejected")
```


4. Create a map of North Carolina where the counties have a fill color 
corresponding to the percentage of votes that were cast by mail. Carefully
consider effective labels, colorings, and aesthetics.

```{r ex-4}
ggplot(nc_votes) +
  geom_sf(aes(fill = mailed/(mailed + voted))) +
  scale_fill_gradient(low = "#fff7f3", high = "#49006a") +
  theme_bw() +
  labs(fill = "Proportion of votes cast by mail")
```


5. Explore on your own and create an effective spatial data visualization using
these data.

```{r ex-5}
ggplot(nc_votes) +
  geom_sf(aes(fill = (voted + mailed) / (regstrd))) +
  scale_fill_gradient(low = "#fff7f3", high = "#49006a") +
  theme_bw() +
  labs(fill = "Proportion",
       title = "Proportion of Registered Voters who Cast Ballot",
       subtitle = "By NC County")
```

