---
title: "AE 08"
subtitle: "For Loops and the Bootstrap"
date: "Due: Sep 16, 11:59 PM"
author: "Scott Burstein"
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

## Clone a repo + start a new project

Go to the `ae08-[GITHUB USERNAME]` repo, clone it, and start a new project in 
RStudio. See the earlier labs and application exercises for detailed instructions.

## Configure git

Run the following code to configure Git. Fill in your GitHub username and the 
email address associated with your GitHub account. 

```{r eval = FALSE}
library(usethis)
use_git_config(user.name= "sburstein", user.email="scott.burstein@duke.edu")
```

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

# For loops

```{r for-loops}

output <- numeric(10)

for(i in 1:10){   #the sequence

  output[i] <- i        #the body  
  
}
i
```

# General Social Survey

The General Social Survey (GSS) gathers data on American society to "monitor and 
explain trends and constants in attitudes, behaviors, and attributes [and] to 
make high-quality data easily accessible to scholars, students, policy makers, 
and others".

The GSS data includes questions on demographics, behavior, attitudes, and topics 
of special interest. Topics include civil liberties, nationalism, crime, 
religion, national spending, morality, and many others.

The target population is adults living in households in the United States and
data has been collected since 1972. You can read more about the methodology 
[here](https://gss.norc.org).

We will analyze a single question from the 2016 GSS which asked "About how many 
[minutes] per week do you spend sending and answering electronic mail or e-mail?"

Some cleaning and manipulation has already been done. A representative sample of 
114 adult Americans is stored in the `gss_email.csv` file.

```{r read-in-data}
gss_email <- read_csv("data/gss_email.csv")
```

Let's calculate a 95\% confidence interval for the mean amount of time Americans
spend on email weekly.

```{r bootstrap-mean}
boot_dist <- numeric(1000) 

for(i in 1:1000){ 
  set.seed(i) 
  indices <- sample(1:nrow(gss_email), replace = TRUE) 
  
  boot_mean <- gss_email %>% 
    slice(indices) %>%
    summarize(boot_mean = mean(email_minutes)) %>%  
    pull()                       

  boot_dist[i] <- boot_mean            
}
boot_dist
indices
```

```{r confidence-bounds}
quantile(boot_dist, 0.025)
quantile(boot_dist, 0.975)
```

The 95\% confidence interval is (376.9469 , 609.7276). 

What is the correct interpretation of this interval?

There is a 95% probability that the population mean is between the specified interval bounds.

Poll time!

Construct a visualization of the amount of time Americans spend on email per 
week? What do you notice?

```{r visualize-email}
ggplot(data = gss_email,
       mapping = aes(x = email_minutes)) +
  geom_histogram()
```

Let's calculate a 95\% confidence interval for the **median** amount of time 
Americans spend on email weekly. Remember to remove `eval = FALSE` when you are
done coding.


```{r bootstrap-median, eval = FALSE}
boot_dist <- numeric(1000) 

for(i in 1:1000){ 
  
  set.seed(i) 
  indices <- sample(1:nrow(gss_email), replace = TRUE)
  
  boot_median <- gss_email %>% 
    slice(indices) %>%
    summarize(boot_median = median(email_minutes)) %>%  
    pull()                       

  boot_dist[i] <- boot_median            
}

boot_dist
boot_median
```

```{r confidence-interval}
quantile(boot_dist, 0.025)
quantile(boot_dist, 0.975)
```

Provide a careful interpretation of your interval in the context of the problem.

The 95% confidence interval indicates that we can be 95% confident that the real median value for email minutes per week is between these two values (120 and 300).
