---
title: "AE 07"
subtitle: "Alinity COVID-19 Test"
date: "Due: Sep 11, 11:59 PM"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

## Clone a repo + start a new project

Go to the `ae07-[GITHUB USERNAME]` repo, clone it, and start a new project in 
RStudio. See the earlier labs and application exercises for detailed instructions.

## Configure git

Run the following code to configure Git. Fill in your GitHub username and the 
email address associated with your GitHub account. 

```{r eval = FALSE}
library(usethis)
use_git_config(user.name= "sburstein", user.email="scott.burstein@duke.edu")
```

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

## Warm-up

Suppose 30% of students in a class were Pratt students and 70% were 
Trinity students. 10% of Pratt students are vegetarian and 5% of Trinity students are 
vegetarian. 

### Exercise 1

What is the probability that a randomly selected student in this class is 
vegetarian?

```{r ex-1}
#P = P(veg) = (prop. of Pratt veg. * overall prop. of Pratt students) + (prop. of Trinity veg. * overall prop. of Trinity students)
P = (0.3 * 0.1) + (0.7 * 0.05)
P
#P(Vegetarian) = 0.065
```

### Exercise 2

Given that a student is vegetarian, what is the probability that they are in 
Pratt?

$$P(p|v) = \dfrac{P(v|p)P(p)}{P(v)}$$
$$P(p|v) = \dfrac{P(v|p)P(p)}{P(v|p)P(p) + P(v|t)P(t)}$$

```{r ex-2}
#P = P(Pratt | veg) = P(Pratt & Veg) / P(Veg) = (prop. of Pratt veg. * overall prop. of Pratt students) / P(Vegetarian)
P = (0.3 * 0.1) / ((0.3 * 0.1) + (0.7 * 0.05))
P
#P(Pratt | veg) = 0.4615385
```

## The Abbott Alinity Test

The global coronavirus pandemic illustrates the need for accurate testing of 
COVID-19, as its extreme infectivity poses a significant public health threat. 
Due to the time-sensitive nature of the situation, the FDA has enacted emergency 
authorization of a number of serological tests for COVID-19. Full details of 
these tests may be found on its website 
[here](https://www.fda.gov/medical-devices/coronavirus-disease-2019-covid-19-emergency-use-authorizations-medical-devices/eua-authorized-serology-test-performance).

The Abbott Alinity test has an estimated sensitivity of 100% and specificity of 
99%. 

### Exercise 3

Suppose the prevalence of COVID-19 in the general population is 2%. What is 
the positive predictive value of the Abbott Alinity test?

$$P(D|+) = \dfrac{P(+|D)P(D)}{P(+|D)P(D) + P(+|ND)P(ND)}$$
$$P(D|+) = \dfrac{sens \times prev}{sens \times prev + (1 - spec) \times (1 - prev)}$$

```{r ex-3}
prev <- 0.0002
sens <- 1
spec <- .99

prob <- (sens * prev) / ( (sens * prev) + ((1 - spec) * (1 - prev)) )
prob
#Positive predictive value of the Abbott Alinity Test is 0.0196 given a COVID pop. prevalence of 2%
```

Positive and negative predictive values depend on prevalence.

### Exercise 4

In the context of diagnostic testing, would you prefer to have a test with perfect 
sensitivity and good specificity or would you rather have a test with perfect 
specificity and good sensitivity? Explain your answer.

I would rather have tests that yield perfect sensitivity and good specificity. With COVID, I would rather have a test used that yields a higher proportion of false positives than a test that yields a higher proportion of false negatives. There is no "real" consequence (i.e. expensive medical procedure, etc.) associated with a false positive COVID test besides self-quarantining. Whereas with a false negative, individuals risk spreading COVID unknowingly to others.