---
title: "AE 11"
subtitle: "Parkinson's Disease"
date: "Due: October 7, 11:59PM"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

## Clone a repo + start a new project

Go to the `ae11-[GITHUB USERNAME]` repo, clone it, and start a new project in 
RStudio. See the earlier labs and application exercises for detailed 
instructions.

## Configure git

Run the following code to configure Git. Fill in your GitHub username and the 
email address associated with your GitHub account. 

```{r eval = FALSE}
library(usethis)
use_git_config(user.name= "sburstein", user.email="scott.burstein@gmail.com")
```

```{r load-data, eval = FALSE}
library(tidyverse)

parkinsons <- read_csv("data/parkinsons.csv")
head(parkinsons)
#view(parkinsons)
```

### Exercise 1

Construct and interpret a 95% confidence interval for the difference in mean
shimmer (in dB) between the PD and non-PD individuals using a simulation-based
approach.

```{r sim-based-ci}
set.seed(42)

healthy <- parkinsons %>%
  filter(status == "Healthy")

pd <- parkinsons %>%
  filter(status == "PD")

n_sims <- 1000

boot_diffs <- numeric(n_sims)

for (i in 1:n_sims){
  #create indices
  indices_h <- sample(1:nrow(healthy), replace = TRUE)
  indices_p <- sample(1:nrow(pd), replace = TRUE)
  
  #bootstrap est. group means
  temp_h <- healthy %>%
    slice(indices_h) %>%
    summarize(mean_shimmer = mean(shimmer)) %>%
    select(mean_shimmer) %>%
    pull()
  
  temp_p <- pd %>%
    slice(indices_p) %>%
    summarize(mean_shimmer = mean(shimmer)) %>%
    select(mean_shimmer) %>%
    pull()
  
  # diff. means in bootstrap sample
  #                Healthy    PD
  boot_diffs[i] <- temp_h - temp_p
}
```

```{r plot-bootstrap-dist}
boot_diffs <- tibble(diffs = boot_diffs)

ggplot(boot_diffs, aes(x = diffs)) +
  geom_histogram(fill = "skyblue",
                 color = "darkblue") + 
  labs(x = "", y = "")

boot_diffs %>%
  summarize(lower = quantile(diffs, 0.025),
            upper = quantile(diffs, 0.975))
```


### Exercise 2

Repeat Exercise 1 using the CLT.

```{r ex2-CLT}
t.test(shimmer ~ status,
       data = parkinsons,
       mu = 0,
       var.equal = FALSE,
       alternative = "two.sided",
       conf.level = 0.95)
```


### Exercise 3

Is there evidence that there is a difference in shimmer between these two 
groups? Perform a formal statistical hypothesis test using a CLT-based
method. State your null and alternative hypotheses, your p-value, statistical 
decision, and a conclusion in the context of the problem.

$H_o: \mu_{PD} = \mu_H$

$H_a: \mu_{PD} \neq \mu_H$

```{r clt-test}
t.test(shimmer ~ status,
       data = parkinsons, 
       mu = 0,
       var.equal = FALSE,
       alternative = "two.sided",
       conf.level = 0.95)
```

Repeat Exercise 3 using a simulation-based method.

### Same as Ex. 3 CLT test, modifications necessary...

$H_o: \mu_{PD} = \mu_H$

$H_a: \mu_{PD} \neq \mu_H$

```{r sim-clt-test}
t.test(shimmer ~ status,
       data = parkinsons, 
       mu = 0,
       var.equal = FALSE,
       alternative = "two.sided",
       conf.level = 0.95)
```


