---
title: "Lab 06 - Hypothesis Testing"
author: 'Team Name: Scott Burstein, Morgan Pruchniewski, Katie Zhou'
date: "10/02/20"
output:
  pdf_document: default
  html_document:
    df_print: paged
subtitle: due Thu, Oct 8 at 11:59p
---

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
parkinsons <- read_csv("data/parkinsons.csv")
``` 

# Exercise 1

Write out the null and alternative hypotheses for this question in both words 
and symbols.

$H_o: \mu = 25$
Null hypothesis is that the mean HNR in voice recordings for healthy volunteers 
is equal to 25.

$H_a: \mu \neq 25$
Alternative hypothesis is that the mean HNR in voice recordings for healthy
volunteers is not equal to 25.

# Exercise 2

Display a visualization of your simulated null distribution, and describe the 
values that would cause you to reject your null hypothesis. Does our observed 
sample mean lie in this rejection region?

```{r ex2-filter}
healthy <- parkinsons %>%
  filter(status == "Healthy")
```


```{r ex-2_simulated-null-dist}
set.seed(2)
n_sims <- 5000
stat <- numeric(n_sims)
for (i in 1:n_sims) {
  indices <- sample(1:nrow(healthy), replace = T)
  boot_mean <- healthy %>%
    slice(indices) %>%
    summarize(boot_mean = mean(hnr)) %>%
    pull()
  stat[i] <- boot_mean
}

boot_means <- tibble(stat = stat)
```

```{r ex-2-shift}
boot_mean <- boot_means %>%
  summarize(mean = mean(stat)) %>% 
  pull()

boot_means <- boot_means %>% 
  mutate(shifted_means = stat + (25 - boot_mean))
```


```{r ex-2 visualization}
ggplot(data = boot_means, aes(x = shifted_means)) +
  geom_histogram(
    binwidth = 0.1, 
    color = "darkblue", 
    fill = "skyblue") +
  labs(
    x = "Mean HNR", 
    y = "Count") +
  geom_vline(xintercept = mean(boot_means$stat), 
             lwd = 1, color = "red")
```



```{r ex-2-interval}
bootstrap_interval <- boot_means %>%
  summarize(lower = quantile(stat, 0.025),
            upper = quantile(stat, 0.975))
bootstrap_interval
```


A mean HNR value less than 23.72 or greater than 25.66 would lead us to reject
the null hypothesis. Our observed sample mean lies between these upper and lower 
bounds, so it is not within the rejection region.

# Exercise 3

What is your p-value, decision, and conclusion in context of the research 
question?

```{r sample-mean}
boot_means %>%
  summarize(mean(stat))
```


```{r ex-3-p-value}
boot_means %>%
  filter(shifted_means <= 24.68847 |
    shifted_means >= (25 + (25 - 24.68847))) %>%
  summarize(p_value = n() / nrow(boot_means))
```

Our p-value is .5358, which is larger than $\alpha = 0.05$, so we fail to reject the null hypothesis. This means that we lack significant evidence to say that the mean HNR of healthy people is different from 25.

# Exercise 4

Given your conclusion in Exercise 3, which type of error could you possibly 
have made? What would making such an error mean in the context of the research 
question?

Since we failed to reject the null hypothesis, we could have made a Type II error. This would mean that the true mean HNR of healthy patients actually is significantly different from 25. 

# Exercise 5

Write out the null and alternative hypotheses for this question in both words 
and symbols.
$H_o: \mu = 24.7$
Null hypothesis is that the mean HNR in voice recordings for volunteers with Parkinson's Disease
is equal to 24.7.

$H_a: \mu\neq 24.7$
Alternative hypothesis is that the mean HNR in voice recordings for volunteers with Parkinson's Disease
volunteers is not equal to 24.7.


# Exercise 6

What is the distribution of the test statistic under the null hypothesis, the 
test statistic itself, the p-value, decision, and conclusion in context of the 
research question?

```{r ex-6 test-statistic}
pd <- parkinsons %>%
  filter(status == "PD")

t.test(pd$hnr, mu = 24.7, conf.level = 0.95)
```
The distribution of the test statistic is a t-distribution with n-1 degrees of freedom.

The test statistic itself is: the mean HNR in voice recordings among participants with Parkinson's Disease
The p-value is: < 2.2e-16, less than the significance level $\alpha = 0.05$, meaning we reject the null hypothesis.
Our conclusion in context of this question is that the mean HNR in participants with Parkinson's Disease is not equal to 24.7.


# Exercise 7

Given your conclusion in Exercise 2, which type of error could you possibly
have made? What would making such an error mean in the context of the research 
question?

We could have made a Type I error. This would entail that $H_o$ is true even though our analysis supported its rejection.


# Exercise 8

Would you expect a 95% confidence interval computed using the same data to 
contain 24.7 or not? Explain.

No, because the bounds of the confidence interval don't contain 24.7 and its p-value is very small. This means there is a very small chance of 24.7 being the true mean.

# Exercise 9

Write out the null and alternative hypotheses in words.

$H_o: \rho = 0$
The null hypothesis is that the correlation between voice jitter and voice shimmer among healthy volunteers is 0.

$H_a:\rho \neq 0$
The alternative hypothesis is that the correlation between voice jitter and voice shimmer among healthy volunteers is not zero.

# Exercise 10

Display a visualization of your simulated null distribution, and describe the
values that would cause you to reject your null hypothesis. Does our observed 
sample correlation lie in this rejection region?

```{r ex-10-simulation}
set.seed(10)
n_sims <- 5000
stat <- numeric(n_sims)
for (i in 1:n_sims) {
  indices <- sample(1:nrow(healthy), replace = T)
  boot_corr <- healthy %>%
    slice(indices) %>%
    summarize(boot_corr = cor(jitter,shimmer)) %>%
    pull()
  stat[i] <- boot_corr
}

boot_corrs <- tibble(stat = stat)
```

```{r ex-10-shift}
boot_corrs <- boot_corrs %>% 
  mutate(shifted_corrs = stat + (0 - boot_corr))
```

```{r ex-10-visualization}
ggplot(data = boot_corrs, aes(x = shifted_corrs)) +
  geom_histogram(
    binwidth = 0.03, 
    color = "darkblue", 
    fill = "skyblue") +
  labs(
    x = "Correlation Between Voice Jitter and Shimmer", 
    y = "Count") +
  geom_vline(xintercept = mean(boot_corrs$stat), 
             lwd = 1, color = "red")
```
```{r ex-11-test-info}
correlations_interval <- boot_corrs %>%
  summarize(lower = quantile(shifted_corrs, 0.005),
            upper = quantile(shifted_corrs, 0.995))
correlations_interval
```

```{r ex-11-sample-correlation}
boot_corrs %>% 
  summarize(correlation_mean = mean(stat))
```

A sample correlation of under -0.1465 or over 0.2711 would cause us to reject the null hypothesis. Our observed sample correlation is 0.634, which is not within these bounds, so it is within the rejection region.


# Exercise 11

What is your p-value, decision, and conclusion in context of the research 
question?

```{r ex-11-p-value}
boot_corrs %>%
  filter(shifted_corrs >= 0.6340549 |
         shifted_corrs <= - 0.6340549) %>%
  summarize(p_value = n() / nrow(boot_corrs))
```

Our p-value is 0, which is under the significance level $\alpha = 0.01$. We can then reject the null hypothesis that the correlation $\rho = 0$. 

In context of this problem, the p-value of 0 means there is zero probability that the population correlation of voice jitter and voice shimmer is as or more extreme than 0. This means that the voice jitter and voice shimmer of healthy volunteers are correlated.