---
title: "Bicycle Data Wrangling"
author: "Scott Burstein"
date: "`r Sys.Date()`"
output: html_document
---



## Clone a repo + start a new project

Go to the `ae04-[GITHUB USERNAME]` repo, clone it, and start a new project in RStudio. See previous labs and application exercises for more detailed instructions about cloning a repo and starting a project.

### Configure git

Run the following code to configure Git. Fill in your GitHub username and the email address associated with your GitHub account. 

```{r eval=FALSE}
library(usethis)
use_git_config(user.name= "sburstein", user.email="scott.burstein@duke.edu")
```

## Practice with data wrangling

```{r load-packages, warning = FALSE, message = FALSE}
library(tidyverse)
```

The data set contains all North Carolina bike crash data from 2007-2014. The data was originally downloaded on September 6, 2018 and is available in the dsbox R package. 

Use the code below to load the data from the .csv file

```{r load-data, warning = FALSE, message = FALSE}
ncbikecrash <- read_csv("data/ncbikecrash.csv")
```

### Part 1

**Display the first 10 rows of the ncbikecrash data set.**

```{r first-10}
ncbikecrash %>%
  slice(1:10)
```


### Part 2 

**Filter for bike crashes in Durham County where the driver age group is 0 to 19.**

```{r Durham-age-crashes}
ncbikecrash %>%
  filter(driver_age_group == "0-19", county == "Durham")
```

### Part 3

**On which day of the week did the most bike crashes occur? Use the `count` function to get the number of bike crashes for each day (`crash_day`). Display the results in descending order.**

```{r crashes-by-day}
ncbikecrash %>%
  count(crash_day, sort = TRUE)
```

### Part 4

**Calculate the proportion of crashes that are hit and run (`hit_run`) for each hour (`crash_hour`)**


```{r hit-run}
ncbikecrash %>%
  group_by(crash_hour) %>%
  summarize(prop_hitrun = mean(hit_run == "Yes"))
```


### Part 5: Just for fun!

**Construct an effective visualization of the data from part 4**

```{r hit-run-vis}
ncbikecrash %>%
  group_by(crash_hour) %>%
  summarize(prop_hitrun = mean(hit_run == "Yes"),
            count = n()) %>%
  ggplot(mapping = aes(x = crash_hour, y = prop_hitrun)) +
  geom_point(aes(size = count)) +
  labs(title = "Proportion of Hit & Runs by Crash Hour", 
       subtitle = "Point Size Correlates to # of Crashes Per Hour",
       x = "Hour of Day (0-23)",
       y = "Proportion of Crashes that are Hit & Runs")
```