---
title: "Joining data sets + data wrangling"
author: "Scott Burstein"
date: "`r Sys.Date()`"
output: github_document
---

## Clone a repo + start a new project

Go to the `ae05-[GITHUB USERNAME]` repo, clone it, and start a new project in RStudio. See the earlier labs for detailed instructions.

## Configure git

Run the following code to configure Git. Fill in your GitHub username and the email address associated with your GitHub account. 

```{r eval = FALSE}
library(usethis)
use_git_config(user.name= "sburstein", user.email="scott.burstein@duke.edu")
```

```{r setup, include = FALSE}
library(tidyverse)
library(nycflights13)
library(maps)
```

1. Run the setup code chunk. If you receive an error, you will need to run `install.packages("nycflights13")` and `install.packages("maps")` in the console first then re-run the setup code chunk.

2. Familiarize yourself with the datasets `flights` and `airports` using the documentation `?flights`, `glimpse()`, and other functions. What information is provided in each dataset?

```{r familiarization-w-flights}
?flights
glimpse(flights)
```
flights provides information for 336,776 flights across 19 variables.

```{r familiarization-w-airports}
?airports
glimpse(airports)
```
airports provides information for 1,458 airports across 8 variables.

3. Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need from `flights`? What variables would you need from `airports`? Consider (in words) how you might join these two datasets.

In order to create an approximate visualization for the route each plane flies from its origin to its destination, the following information would be necessary:

from `flights`  :  origin & destination (both are <chr> variables)
from `airports` :  lat & lon (both are <dbl> variables)

A potential method for joining these two datasets would be to perform a left_join() on  `flights`, `airports` that retains all information for each flight in `flights`. Next, filter to just retain lat and lon <dbl> variables for each flight's origin and destination <chr> data. Draw straight lines from the origin to the destination and overlay on an appropriately scaled map.

One interesting addition to this visualization would be to color each line by airline carrier to get a sense of what routes and regions each airline flies the most.

### What is the geographic distribution of delays in the United States?

4. Starting with the `flights` dataset, find the average delay by destination and store your results in a dataset named `delays`. Note that NA's are "contagious" so you will need to add an argument to the `mean` function.

```{r average-delay-by-destination, eval = FALSE}
delays <- flights %>%
  group_by(dest) %>%
  summarize(delay = mean(arr_delay, na.rm = TRUE), count=n())
delays
```

5. Create a new dataset by joining `delays` and `airports`. Only include observations that have both delay and airport information. Note `delays` and `flights` have no variables in common so you will need to specify the variables to join using `by`.

```{r delays-inner-join-airports, eval = FALSE}
delays_airports <- delays %>%
  inner_join(airports, by = c("dest" = "faa"))
delays_airports

left_out <- delays %>%
  anti_join(airports, by = c("dest" = "faa"))
left_out
```

6. Create a map to show the spatial distribution of delays. Some helper code is provided. Use an appropriate aesthetic to display the average delay for each airport.

```{r Spatial-Distribution-of-Delays, eval = FALSE}
delays_airports %>%
  ggplot(aes(x = lon, y = lat)) +
    borders("state") +
    geom_point(aes(color = delay)) +
    coord_quickmap()
```

### Is there a relationship between the age of a plane and its delays? The plane tail number is given in the `tailnum` variable in the dataset `flights`. The year the plane was manufactured is given in the `year` variable in the dataset `planes`.

7. Start by finding the average arrival delay for each plane and store the resulting dataset in `plane_delays`.

```{r average-arrival-delay}

#plane_delays is average arrival delay for each plane

plane_delays <- flights %>%
  select(tailnum, arr_delay) %>%
  group_by(tailnum) %>%
  summarise(avg_arr_delay = sum(arr_delay, na.rm = TRUE) / length(arr_delay))

plane_delays
```

8. Join `plane_delays` to the `planes` data using an appropriate join and then use `mutate` to create an `age` variable. Note this data is from 2013.

```{r join}
LJ_delays_planes <- left_join(plane_delays, planes)

# Performed a left join on plane_delays, planes. data joined based on tailnum.

view(LJ_delays_planes)
```

```{r age-variable}

#mutate to find age from manufacture age
#Only determining plane age based on year of plane commission.
#All flights from dataset occurred in 2013

plane_age <- LJ_delays_planes %>%
  mutate(age = 2013 - year)

view(plane_age)
```

```{r cleaning-dataset}

#just keep two useful columns (aircraft age and average arrival delay)

age_avgDelay <- plane_age %>%
  select(age, avg_arr_delay) %>%
  group_by(age) %>%
  na.omit() %>%
  summarise(avg_arr_delay_by_age = sum(avg_arr_delay) / length(avg_arr_delay))

age_avgDelay
```


9. Finally, create an effective visualization of the data.

```{r Average Delay Visualization Plot}

ggplot(data = age_avgDelay,
       mapping = aes(x = age, y = avg_arr_delay_by_age)) +
  labs(title = "Average Arrival Delay by Aircraft Age",
       x = "Aircraft Age",
       y = "Average Arrival Delay") +
  geom_bar(stat = "identity")

```

I ultimately decided to create a bar plot visualization. There are some outliers which are the result of small sample size n for aircrafts of age greater than ~35 years. Interesting takeaway is that average aircraft delay is not explicitly correlated with aircraft age.
