---
title: "AE 13"
subtitle: "Car Prices"
date: "Due: October 16, 11:59PM"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

```{r load-packages, include = FALSE}
library(tidyverse)
library(broom)
```

The dataset `car_prices.csv` contains attributes of cars offered for sale on 
cars.com. The data was collected in 2017 and is taken from the `Stat2Data` 
package.

- `type`: Model (Accord, Maxima, Mazda6)
- `age`: Age of the used car (in years)
- `price`: Price (in thousands of dollars)
- `mileage`: Previous miles driven (in thousands of miles)

```{r load-data, message = FALSE}
car_prices <- read_csv("data/car_prices.csv")
```


## Price and Mileage

Consider a regression model with the response `price` and a single predictor 
`mileage`. 

1. Create a scatterplot of price and mileage.

```{r scatterplot}
ggplot(car_prices,
       aes(x = mileage, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

```


2. Use appropriate functions in `R` to find the fitted model and display the 
results in tidy format. Write out the equation of the fitted model and visualize
the linear model on your scatterplot.

```{r linear-model}
m_pr_mile <- lm(price ~ mileage, data = car_prices)
tidy(m_pr_mile) %>%
  select(term, estimate)
```
$$\widehat{Price} = 21.4 - 0.128 \times Mileage$$

3. Carefully interpret the slope and intercept in the context of the problem.

- **Slope**: For every additional thousand miles on the used car, the price is 
expected to decrease, on average, by $128.

- **Intercept**: Used cars that have 0 miles on them are expected to cost 
$21,400, on average.

4. What is the predicted selling price of a car with 50,000 miles?

```{r pred}
21.4 - 0.128 * 50
```


5. Suppose my aging Accord has 225,000 miles. Is it appropriate to use this 
model to make a prediction for the selling price? Why or why not?

## Price and Model

No, because 225,000 is beyond range of data and the curve appears to flatten out near the end of x-bound (upper end).

Consider a regression model with the response `price` and the categorical 
predictor `type` (Accord, Maxima, Mazda6).

6. Use appropriate functions in `R` to find the fitted model and display the 
results in tidy format. Write out the equation of the fitted model.

```{r fitted-model-car-type}
m_pr_type <- lm(price ~ type, data = car_prices)
tidy(m_pr_type) %>%
  select(term, estimate)
```

$$\widehat{Price} = 14.3 + 1.19 \times typeMaxima - 2.77 \times type Mazda6$$
Maxima:
$$\widehat{Price} = 14.3 + 1.19$$
Mazda:
$$\widehat{Price} = 14.3 - 2.77$$
Honda:
$$\widehat{Price} = 14.3$$

7. Carefully interpret the intercept and slope(s) in the context of the problem.

- **Slope Maxima**: Maximas are expected, on average, to cost $1,187 more than 
Accords.

- **Slope Mazda6**: Mazda6's are expected, on average, to cost $2,773 less than 
Accords.

- **Intercept**: Accords are expected to cost $14,277

8. How is the categorical variable `type` included in the model?

As two indicator variables, with Accords as the reference level. More generally, 
for a categorical variable with $k$ levels, we include $k-1$ indicator 
variables.

9. What are possible limitations of your two regression models?

- Possible concerns about non-linearity.
- We fit separate models, but really we should fit a single model that includes 
**both** mileage and type as predictors.
- Is this data representative? If not a random sample, we can't generalize.
- What happens when a car is sold?
