---
title: "AE 10"
subtitle: "Data Ethics & Bone density"
date: "Due: September 30, 11:59PM"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

## Data Ethics Module

For today's tasks you determined (part of) what Google and Facebook know about 
you. You are welcome (but not obligated!) to share these results with your
breakout room / the class.

Discuss the questions below with your breakout rooms and write down your 
thoughts in the spaces provided.

### Discussion Question 1

What do Google and Facebook know about you? How accurate is the information?
Were you surprised by any of the results?

Google and Facebook both keep track of my interests through my usage of their
respective apps. Google, for example, states that "Ads are based on personal 
info you've added to your Google Account, data from advertisers that partner 
with Google, and Google's estimation of your interests". It was surprising to 
see a list of at least 100 topics that Google has recorded me having a 
preference towards. Many, of them, I agree with, like "Auto Racing" and 
"Comedy Films" for example. However it was surprising to see some other topics,
like "Fast Food" and "Greeting Cards", which I never would have labeled as 
some of my interests. I definitely want to learn more about how Google and 
Youtube tracks my information and behavior across its domain.

Facebook has a similar page designed to keep track of my interests based on 
advertisements I engage with and Facebook pages I visit. Facebook divides the 
data into more detailed categories, such as "Business and industry" and "News 
and entertainment".

### Discussion Question 2

Were you aware this information was being collected? How do you feel about this?
Relate these results and your feelings to our data ethics discussion last week.

I was aware that this information was being collected. I feel fairly comfortable 
with this reality, however it is alarming to know how detailed the information 
which Google and Facebook keep about me is. I also believe they keep much more 
data that is perhaps hidden or harder for me as the user to view. The consistent 
breach of my digital privacy is unfortunately the price I am willing to pay for 
free access to the services offered by Google and Facebook.

### Discussion Question 3

On the lecture slides there are ad personalization results for two individuals. 
What can you tell me about their backgrounds basely solely on this information?
Who is each person?

The first is a male in his 40s, and the second is a female from St. Paul, 
Minnesota. They also have different interests. Based on this, I infer that the 
first set represents my professor, Dr. Bob Eisinger, and the second represents 
his wife.

## Normal Distribution

Suppose the mean bone density for a 65-year old white woman is normally 
distributed with mean 809 mg/cm3 and standard deviation of 140 mg/cm3 (these are 
actually pretty accurate numbers).

### Exercise 1

What bone densities correspond to the 10th, 50th, and 99th percentiles of
this distribution?

```{r ex-1}
qnorm(.1, mean = 809, sd = 140)
qnorm(.5, mean = 809, sd = 140) #0.5 represents mean
qnorm(.99, mean = 809, sd = 140)
```

### Exercise 2

The densities of two different types of woods are below:

- Plywood: 540 mg/cubic centimeter
- Mahogany: 710 mg/cubic centimeter

What is the probability that a random 65-year old white woman has bones less
dense than each of these woods?

```{r ex-2}
#prob less dense than Plywood
pnorm(540, mean = 809, sd = 140)
#prob less dense than Mahogany
pnorm(710, mean = 809, sd = 140, lower.tail = FALSE)

### (VALUE - AVERAGE) / SD
z <- (540 - 809) / 140
pnorm(z)
```

### Exercise 3

What is the probability that, for a *group* of 10 randomly sampled 65-year old
white women, the mean bone density of the *group* is less dense than these woods?

Can use statistics even without knowledge of sample size since normal 
distribution properties still apply.

```{r ex-3}

population_mean <- 809
standard_error <- 140 / sqrt(10)

pnorm(540, mean = population_mean, sd = standard_error)

# SAMPLE MEAN ~ N(MU, SIGMA / SQRT(N))

#prob group density less than Plywood
pnorm(540, mean = 809, sd = 140 / sqrt(10))
#prob group density less than Mahogany
pnorm(710, mean = 809, sd = 140 / sqrt(10))
```

Large N -> Smaller spread of sample mean

### Exercise 4

Explain the differences between your answers in Exercise 2 and 3.

The probabilities in ex-3 are substantially smaller than the probabilities in 
ex-2, which makes sense since group statistics sample means are normally 
distributed (individual measures are as well), however the variability in the 
group values are far less than the variability of single observations.

Less variability in sample mean when take a sample size of 10 rather than 1. Are 
dealing with distribution of sample means, which has normal distribution 
properties according to the Central Limit Theorem.
