---
title: "AE 15"
subtitle: "It's Jeopardy!"
date: "Due: October 30, 11:59PM"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

```{r packages, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(stringr)
```

We will work with a dataset containing the category, value, question, and 
answer for two recent seasons of Jeopardy.

If you've never seen Jeopardy before, watch a brief clip 
[here](https://www.youtube.com/watch?v=qY_q99iaWGw).

- `category`: category of question
- `value`: value of question in dollars
- `question`: text of question
- `answer`: question answer

```{r load-data, message = FALSE}
jeopardy <- read_csv("data/questions.csv")
```

```{r view}
glimpse(jeopardy)
```


1. Write a short line of code that returns all rows where the category has a 
period.

```{r ex-1}
jeopardy %>% 
  filter(str_detect(category, "\\.")) %>% 
  print(n = Inf)
```

2. Write a short line of code that returns all rows where the answer contains 
"North Carolina".

```{r ex-2}
jeopardy %>% 
  filter(str_detect(answer, "North Carolina"))
```


3. Using a single code pipeline, return all rows where the question contains a 
(numeric) year between 1800 and 1999.

```{r ex-3}
jeopardy %>% 
  filter(str_detect(question, "1[89]\\d\\d")) %>% #"\\d\\d\\d\\d" all 4-digit numbers
  print(n = Inf)

```


4. Use `str_sub` to extract the **last three** characters from every answer 
and store the results in a new column. Do not use `str_length`.

```{r ex-4}
jeopardy %>% 
  mutate(last_three = str_sub(answer, -3, -1)) %>% 
  select(answer, last_three)
```


5. Using a single code pipeline, return all rows with answers that end with 
three vowels.

```{r ex-5}
jeopardy %>% 
  filter(str_detect(answer, "[AEIOUaeiou][AEIOUaeiou][AEIOUaeiou]$")) %>% 
  select(answer)
```


6. Using a single code pipeline, return all answers that end with `ng` but not
`ing`.

```{r ex-6}
jeopardy %>% 
  filter(str_detect(answer, "[^i]ng$")) %>% 
  select(answer)
```


7. Use a single code pipeline to create a new variable `prop_vowel` that
is the proportion of all letters in each answer that are vowels. What is the
highest? Lowest?

```{r ex-7}
jeopardy %>% 
  mutate(vowels = str_count(answer, "[AEIOUaeiou]"),
         letters = str_count(answer, "[[:alpha:]]"),
         prop_vowel = vowels / letters) %>%
  arrange(desc(prop_vowel)) %>% 
  select(answer, vowels, letters, prop_vowel) %>% 
  filter(!is.na(prop_vowel), letters > 6) %>% 
  slice(1:3, n())
```


