---
title: "Lab 05 - Bootstrap estimation"
subtitle: "due Thu, Sep 24 at 11:59p"
author: "The Outliers: Morgan Pruchniewski, Scott Burstein, Katie Zhou"
date: "09/18/20"
output:
  html_document: default
  pdf_document: default
---

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
durham_survey <- read_csv("data/durham_survey.csv")
view(durham_survey)
``` 

# Exercise 1

Provide a point estimate of the mean satisfaction with the fire department 
(`quality_fire`) among Durham residents in 2018.

```{r ex-1_mean-satisfaction-fire-dept}
mean_satisfaction_fire_dept <- durham_survey %>%
  filter(quality_fire != 9) %>%
  summarize(mean = mean(quality_fire)) %>%
  print()
```

# Exercise 2

Construct a 95% bootstrap interval for the mean satisfaction with
the fire department among durham residents in 2018. Use at least 1,000 bootstrap
samples. Make sure your interval is *reproducible*.

```{r ex-2_95-bootstrap-int}

boot_dist <- numeric(1000)

for (i in 1 : 1000) {
  set.seed(i)
  indices <- sample(1 : nrow(durham_survey), replace = TRUE)
  boot_mean <- durham_survey %>%
    slice(indices) %>%
    filter(quality_fire != 9) %>%
    summarize(boot_mean = mean(quality_fire)) %>%
    pull()
  boot_dist[i] <- boot_mean
}

boot_means <- tibble(boot_dist)

bootstrap_interval <- boot_means %>%
  summarize(lower = quantile(boot_dist, 0.025),
            upper = quantile(boot_dist, 0.975)
            )
bootstrap_interval
```

# Exercise 3

Visualize the bootstrap distribution and your confidence interval from 
Exercise 2 on a single plot. Remember to include an informative title and axis 
labels. Provide a careful interpretation of your interval in the context of the 
problem.

If you need help visualizing your bootstrap distribution and confidence interval on 
the same plot, check out slides 35 - 40 from Tuesday's bootstrap lecture. This walks 
through creating the plot step-by-step. Your final plot for exercise 3, 6, and 9 
should look similar to the plot on slide 40.

```{r ex-3-bootstrap-distribution-visual}
ggplot(data = boot_means, aes(x = boot_dist)) +
  geom_histogram(color = "black", fill = "blue", alpha = .8) + 
  geom_vline(xintercept = c(bootstrap_interval$lower, 
                            bootstrap_interval$upper),
             color = "red") +
  labs(title = "Bootstrap Distribution of Means for Durham Fire Quality",
       x = "Bootstrap Means", y = "Count")
```

If we computed 1000 samples of the mean for Durham's fire quality satisfaction, we can be 95% confident that the actual mean is within the specified bounds.

# Exercise 4

Provide a point estimate of the *proportion* of the respondents in the 
survey who were satisfied (score of 4 or 5) with the quality of parks and
recreation (`quality_parks_rec`) in Durham. 

```{r parks-rec-satisfaction}
satisfied_parks_rec <- durham_survey %>%
  mutate(satisfied = quality_parks_rec == 4 | quality_parks_rec == 5) %>%
  summarize(proportion_satisfied = mean(satisfied)) %>% 
  print()
```
0.5477 of Durham residents responded that they were satisfied with the quality of parks and recreation.


# Exercise 5

Construct a **99%** two-sided bootstrap interval for the proportion of 
respondents in the survey who were satisfied with the quality of parks and 
recreation in Durham. Use at least 1,000 bootstrap samples. Make sure your
interval is *reproducible*.

```{r ex-5_99-bootstrap-int}
boot_dist <- numeric(1000)

for (i in 1 : 1000) {
  set.seed(i)
  indices <- sample(1 : nrow(durham_survey), replace = TRUE)
  boot_prop_satisfied <- durham_survey %>%
  slice(indices) %>%
  mutate(satisfied = quality_parks_rec == 4 | quality_parks_rec == 5) %>%
  summarize(proportion_satisfied = mean(satisfied)) %>%
    pull()
  boot_dist[i] <- boot_prop_satisfied
}

boot_props <- tibble(boot_dist)

bootstrap_interval <- boot_props %>%
  summarize(lower = quantile(boot_dist, 0.005),
            upper = quantile(boot_dist, 0.995)
            )
bootstrap_interval
#need to remove 9 values from mean prop.
```

# Exercise 6

Visualize the bootstrap distribution and your confidence interval from 
Exercise 5 on a single plot. Remember to include an informative title and axis
labels. Interpret the confidence interval you constructed.

```{r parks-bootstrap-visualization}
ggplot(data = boot_props, aes(x = boot_dist)) +
  geom_histogram(color = "black", fill = "blue", alpha = .8) + 
  geom_vline(xintercept = c(bootstrap_interval$lower, 
                            bootstrap_interval$upper),
             color = "red") +
  labs(title = "Bootstrap Distribution of Satisfaction for Parks/Recreation",
       x = "Bootstrap Proportion Satisfied/Unsatisfied", y = "Count")
```

If we computed 1000 samples of the given data, 99% of those simulations within the marked interval would contain the true proportion of satisfaction for the Parks/Recreation department. 



# Exercise 7

Provide a point estimate of the *correlation* between survey scores of
perceived quality of bike paths (`quality_bike_path`) and the perceived quality 
of pedestrian paths (`quality_ped_path`). 

```{r ex-7_correlation}
correlation_paths <- durham_survey %>%
  filter(quality_bike_path != 9,
         quality_ped_path != 9) %>%
  summarize(r_value = cor(quality_bike_path, quality_ped_path)) %>%
  print()
```

# Exercise 8

Construct a 95% bootstrap confidence interval for the correlation between perceived 
quality of bike paths and pedestrian paths. Use at least 1,000 bootstrap samples. 
Make sure your interval is *reproducible*.

```{r ex-8_correlation_bootstrap}
boot_dist <- numeric(1000)

for (i in 1 : 1000) {
  set.seed(i)
  indices <- sample(1 : nrow(durham_survey), replace = TRUE)
  boot_correlation <- durham_survey %>%
  slice(indices) %>%
  filter(quality_bike_path != 9,
         quality_ped_path != 9) %>%
  summarize(r_value = cor(quality_bike_path, quality_ped_path)) %>%
    pull()
  boot_dist[i] <- boot_correlation
}

boot_cor <- tibble(boot_dist)

bootstrap_interval <- boot_cor %>%
  summarize(lower = quantile(boot_dist, 0.025),
            upper = quantile(boot_dist, 0.975)
            )
bootstrap_interval
```

# Exercise 9

Visualize the bootstrap distribution and your confidence interval from
Exercise 8. Remember to include an informative title and axis labels. Interpret
the confidence interval you constructed.

```{r ex-9 corr-bootstrap-visualization}
ggplot(data = boot_cor, aes(x = boot_dist)) +
  geom_histogram(color = "black", fill = "blue", alpha = .8) + 
  geom_vline(xintercept = c(bootstrap_interval$lower, 
                            bootstrap_interval$upper),
             color = "red") +
  labs(title = "Bootstrap Distribution of Correlation Between Percieved Quality",
       subtitle =  "of Pedestrian Paths and Bike Paths",
       x = "bootstrap correlation between ped. path and bike path satisfaction", y = "Count")
```
If we computed 1000 samples of the given data, 95% of those simulations within the marked interval would contain the true correlation of between pedestrian path satisfaction and bike path satisfaction.