---
title: "In-Class Notes 10-22"
subtitle: "Cherry Trees"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

```{r packages, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(broom)
```

A silviculturist is interested in finding the volume of black cherry trees in 
order to determine timber yield. This is difficult to measure, so the researcher 
uses multiple regression to predict volume (cubic feet) based on height (feet) 
and girth (inches) for a random sample of 31 trees that have been cut down.

Both height and girth are easy to measure so the idea is one can hopefully 
have a reasonable prediction for volume based on girth and height. Data 
taken from *OpenIntro Statistics (4th edition)*.

```{r load-data, message = FALSE}
trees <- read_csv("data/trees.csv")
head(trees)
nrow(trees)
glimpse(trees)
```

## Multiple Regression

1. Fit a multiple linear regression model predicting `Volume` based on `Height`
and `Girth` and display the R output in tidy format.

```{r ex-1}
ml <- lm(Volume ~ Height + Girth, data = trees)
tidy(ml) %>% 
  select(term, estimate) %>% 
  mutate(estimate = round(estimate, 2))
```

2. Write out the equation of the estimated regression line.

$$\widehat{volume} = -58 + 0.34~Height + 4.71~Girth$$

3. Create a 95% confidence interval for the population slope for girth and 
provide an interpretation in the context of the problem.

```{r ex-3}
tidy(ml, conf.int = TRUE, conf.level = 0.95) %>% 
  filter(term == "Girth") %>% 
  select(term, conf.low, conf.high)
```

We are 95% confident that for a one inch increase in girth, volume is predicted 
to increase, on average, between 4.17 and 5.24 cubic inches holding all else 
constant.

4. Is there evidence of a relationship between height and volume? Answer 
formally using a CLT-based hypothesis test.

$H_o: \beta_{Height} = 0$
$H_a: \beta_{height} \neq 0$

The test statistic is 2.6 and the $p$-value is less than 0.001 so we reject the 
null hypothesis. The data provides convincing evidence that height and volume 
are related. We have significant evidence that the true slope is different from 
zero.

5. Consider your fitted model from Exercise 1. Are the conditions for linear 
regression met? Check each condition by constructing appropriate diagnostic 
plots and commenting on each.

```{r ex-5}
m1_aug <- augment(m1)
m1_aug
```

```{r diagnostics}
# volume versus height
ggplot(m1_aug,
       aes(y = Volume, x = Height)) +
  geom_point()

# volume versus girth
ggplot(m1_aug, 
       aes(y = Volume, x = Girth)) +
  geom_point()

# residuals in order of collection
ggplot(m1_aug,
       aes(y = .resid, x = 1:nrow(m1_aug))) +
  geom_point()

# residuals versus fitted
ggplot(m1_aug,
       aes(y = .resid, x = .fitted)) +
  geom_point() +
  geom_hline(yintercept = 0, lwd = 2, lty = 2, col = "red")

#normality histogram
ggplot(m1_aug,
       aes(x = .resid)) +
  geom_histogram(bins = 15)

# normality qq line
ggplot(m1_aug, mapping = aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line()
```




