---
title: "AE 12"
subtitle: "Political Testing"
date: "Due: October 9, 11:59PM"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

## Clone a repo + start a new project

Go to the `ae12-[GITHUB USERNAME]` repo, clone it, and start a new project in 
RStudio. See the earlier labs and application exercises for detailed 
instructions. 

Also remember to configure git.

```{r load-packges, message = FALSE}
library(tidyverse)
```

## Sex and Political Affiliation

Knit your document to html to see the table below, which contains data from the 
2000 General Social Survey.  Sex and political party affiliation are 
cross-classified and you would like to test for independence of sex and party 
affiliation. Use R as a calculator to answer the following questions.

|           | Democrat | Independent | Republican | Total |
| --------- | -------- | ----------- | ---------- | ----- |
| Female    | 762      | 327         | 468        | 1557  |
| Male      | 484      | 239         | 477        | 1200  |
| Total     | 1246     | 566         | 945        | 2757  |

1. What is the expected count of Republican males under the null hypothesis of 
independence?

```{r ex-1}
#expected republican count:
(945 / 2757) * (1200 / 2757) * 2757

#alternative method:
(1200 * 945) / 2757
```


2. What is the contribution of the Republican male cell to the test statistic? 

```{r ex-2}
observed <- 477
expected <- (1200 / 2757) * 945

#chi-squared test statistic
(observed - expected)^2 / expected
```

If chi-squared is 0, summary statistic is what was expected based on observed. 
If large enough, sufficient to reject null hypothesis.

3. What is the distribution of the test statistic? 

$\chi^2$ with 2 degrees of freedom.

4. Suppose the test statistic is 30.1. Use the `pchisq` function to determine 
the p-value and statistical decision.

```{r ex-4}

pchisq(30.1, df = 2)

#correct p-values
1 - pchisq(30.1, df = 2)

pchisq(30.1, df = 2, lower.tail = FALSE)

```
As or more favorable than null hypothesis is small region to right. These are 
bigger values than what was observed.

## Direction of Country

Economist / YouGov surveyed 1,115 U.S. adult citizens between November 3 and 
November 5, 2019.  They asked "Would you say things in this country today are 
generally headed in the right direction, or are they off on the wrong track?" 
and also asked the respondent to report their family income.

```{r read-in-data}
direction_of_country <- read_csv("data/direction.csv")
```

The file `direction.csv` contains each each individual's family income and 
response to this question.

5. Summarize your data in a table and comment on the relationship, reporting 
numerical summaries. What is going on here?

```{r summarize}
table(direction_of_country$income, direction_of_country$direction)
```

6. State the appropriate null and alternative hypotheses.

$H_o:$ Income and opinion about direction of the country are independent.

$H_a:$ Income and opinion about direction of the country are not independent. 

7. Use an appropriate function to find the table of expected counts. Hint: Look 
at the documentation.

```{r ex-7}
x <- table(direction_of_country$income, direction_of_country$direction)

chisq.test(x)$expected
chisq.test(x)$observed

chisq.test(x)
```


8. Use an appropriate function to find the $\chi^2$ test statistic.

```{r ex-8}
chisq.test(x)
```


9. Report your p-value and a conclusion in the context of the problem.

We reject the null hypothesis of independence. We have convincing evidence to 
suggest that family income and opinion about direction of the country are not 
independent. Our test statistic is 34.994 and our $p$-value is less than 0.001.

10. Comment on any limitations of this analysis.

Rows of data categories are not ordered by nominal value. The data is not 
structured according to numeric properties of data, but rather split into 
arbitrary categories and not ordered with valid relationships.

