---
title: "HW 03 - Kid's Growth"
subtitle: "due Mon, Nov 9 at 11:59p"
author: "Scott Burstein"
date: ""
output:
  pdf_document: default
  html_document: default
---

# Configure Git

```{r eval = FALSE}
library(usethis)
use_git_config(user.name= "sburstein", user.email="scott.burstein@duke.edu")
```

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
library(broom)
``` 

```{r read-in-data, message = FALSE}
kids <- read_csv("data/kids.csv")
```

### Exercise 1

```{r ex-1}
kids$sex <- kids$sex %>% 
  factor(labels = c("male", "female"))
```

### Exercise 2

```{r ex-2}
m_ht_sexAge <- lm(height ~ factor(sex) + age, data = kids)
tidy(m_ht_sexAge) %>% 
  select(term, estimate) %>% 
  mutate(estimate = round(estimate, 5))
```

$$\widehat{height} = 87.40 + 0.43~age - 1.16~female$$

### Exercise 3

- **Intercept**: Male newborns (age = 0) are expected to have a birth height of 
87.40 cm. This does not make sense, since the average length for newborns is 
about 50 cm according to Healthline, which is significantly less than the linear 
model indicates. Source Cited:
https://www.healthline.com/health/parenting/average-baby-length#:~:text=
The%20average%20length%20at%20birth,(45.7%20to%2060%20cm).

- **age**: For every one month increase in age, the height of a child
is predicted to increase, on average, 0.43 cm holding all else constant. This 
makes sense given the context of the data and problem.

- **female**: Female children are predicted to have a height that is 1.16 cm 
less than male children, holding all else constant. 
This makes some sense, 
however it is true that females typically grow taller earlier on in their 
development than males, who usually grow, on average, to be taller by the time 
they are fully developed. This being said, it would make more sense to have an 
age variable coefficient that is also dependent on gender.

### Exercise 4

Construct a 90% confidence interval for $\beta_{age}$:

```{r ex-4}
tidy(m_ht_sexAge, conf.int = TRUE, conf.level = 0.90) %>% 
  filter(term == "age")
age_child <- lm(height ~ age, data = kids)
tidy(age_child)
```

We are 90% confident that for a one month increase in the child's age, the 
height is predicted to increase, on average, between 0.418 and 0.448 
centimeters, holding all else constant.

### Exercise 5

```{r ex-5}
glance(m_ht_sexAge) %>% 
  pull(r.squared)
```

$R^2$ = 0.6951797

Approximately 69.5% of the variability in children's height can be explained by 
variation in age and gender

### Exercise 6

```{r ex-6}
age_months = 14.5 * 12
pred_14_male <- 87.3996177 + 0.4325439 * age_months - 1.1615899 * 0 
pred_14_male
```

The predicted height for a fourteen-and-a-half-year-old male is 162.66 cm.

### Exercise 7

Scatterplot of height versus age with points colored by sex:

```{r ex-7}
predicted_vals <- predict(m_ht_sexAge)
height_ageSex_scatter <- ggplot(data = kids,
                                aes(x = age, y = height, color = sex)) +
                            geom_point(alpha = 0.5) +
                            geom_line(aes(y = predicted_vals))
                            
height_ageSex_scatter
```

### Exercise 8

```{r ex-8}
m_ht_sexAge_interact <- lm(height ~ sex + age + sex*age, data = kids)
tidy(m_ht_sexAge_interact) %>% 
  select(term, estimate) %>% 
  mutate(estimate = round(estimate, 5))
```

$$\widehat{height} = 81.35 + 0.47~age + 10.76~female - 0.08~ age \times female$$

### Exercise 9

**males:**
$$\widehat{height_{male}} = 81.35 + 0.47~age$$

**females:**
$$\widehat{height_{female}} = 92.11 + 0.39~age$$
#female intercept = 92.7

### Exercise 10

$H_o:$
Null hypothesis is that the relationship between age and height does not depend 
on sex.

The linear model's slope (rate of change) for height is not affected by the 
interaction between age and sex. This is what the linear model from Exercise 2 
shows.

$H_a:$
Alternative hypothesis is that the relationship between age and height does
depend on sex.

The linear model's slope (rate of change) for height is affected by the 
interaction between age and sex. This is what the linear model from exercise 8 
shows.

```{r ex-10}

#p-value for null hypothesis model:
tidy(m_ht_sexAge)
#equals ~0

#p-value for alternative hypothesis model:
tidy(m_ht_sexAge_interact)
#also equals ~0
```

Since p-value for the null hypothesis (first tbl_df) is effectively 0 
(less than 0.05), there is sufficient evidence to reject the null hypothesis 
that the relationship between age and height does not depend on sex.

Thus, it is most reasonable to conclude that the relationship between age and 
height does depend on sex.


### Exericse 11

```{r ex-11}

predicted_vals_interact <- predict(m_ht_sexAge_interact)
height_ageSex_codep_scatter <- ggplot(data = kids,
                                aes(x = age, y = height, color = sex)) +
                            geom_point(alpha = 0.5) +
                            geom_line(aes(y=predicted_vals_interact)) 
                           #geom_smooth(method = "lm", se = FALSE)
                            
height_ageSex_codep_scatter
```

### Exercise 12


#### Tidy Output Estimates for "Interact" Model:

```{r ex-12_tidy}
tidy(m_ht_sexAge_interact) %>%
select(term, estimate) %>% 
mutate(estimate = round(estimate, 5))
```

Estimates for the Tidy Output Model with `sex`*`age` interaction shown above.

#### "Interact" Model Agument Output:

```{r ex-12_augment}
m_interact_aug <- m_ht_sexAge_interact %>% 
  augment() 
m_interact_aug
```

Estimates for the Augmented Output Model shown above.

#### Residual Plot with Horizontal Line (y = 0):

```{r ex-12_viz_w_line}
indep_viz_w_line <- ggplot(data = m_interact_aug, 
                          aes(x = 1:nrow(kids), y = .resid)) + 
                  geom_point() +
                  labs(x = "Index", 
                       y = "Residual") +
                  geom_hline(yintercept = 0, 
                             lwd = 1.5, 
                             col = "red", 
                             lty = 2) + 
                  labs(x = "Predicted Price", 
                       y = "Residuals")
indep_viz_w_line
```

This residual plot shows that the residuals are random. There is no clear trend 
in the residuals which also is evidence of independence and equal variance.

#### Residual Normality Distribution Plot:

```{r ex-12_normality_hist}
ggplot(m_interact_aug, 
       mapping = aes(x = .resid)) + 
      geom_histogram(bins = 30) +
      labs(x = "Residuals")
```

Shows that the residuals are normally distributed. Strong evidence for 
normality condition being met.

#### Theoretical qq Line Plot:

```{r ex-12_qq_line_plot}
ggplot(m_interact_aug, 
       mapping = aes(sample = .resid)) + 
      stat_qq() +
      stat_qq_line()
```

The quantile-quantile plot confirms that the data is normally distributed.

### Exercise 13

There are some limitations of this model (for predicting the height of 0-16 
year olds based on the data) that is beyond the scope of the diagnostic plots in 
exercise 12.
For example, it is somewhat unreasonable to make conclusions about the growth 
rates for 0-16 year olds when the data has only sampled 8-16 year olds.

Additionally, it might not be true that the relationship between age and height 
is linear (growth rate likely changes with age). This being said, it is 
plausible that some of the model's other conditions may not be met yet either,
such as independence, normality and equal variance. However, based on the 
visualizations in exercise 12, we can be reasonably confident that these 
conditions are satisfied.
