---
title: "AE 17"
subtitle: "Functions"
date: "Due: November 13, 11:59PM"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(rvest)
```

## Functions

1. Rewrite this code snippet as an appropriately named function: 
`x / sum(x, na.rm = TRUE)`.  This code creates weights which sum to 1.

```{r new weights function}
create_weights <- function(x) {
  x / sum(x, na.rm = TRUE)
}

t <- 1:10
create_weights(t)
create_weights(c(1:10, NA))
```


2. Create a function to calculate the standard error of a variable, where SE = 
standard deviation divided by the square root of the sample size. Start with a
vector like `x <- 0:5` and write code to find the SE of x, then turn it into a 
function to handle any vector `x`.

```{r standard error}

se <- function(x) {
  sd(x) / sqrt(length(x))
}

x <- 0:5
se(c(1:100, NA))

```


3. Read the source code for each of the following three functions, puzzle out 
what they do, and then brainstorm better names.

```{r, eval = FALSE}

# Checks if string is matching first letters substr of length prefix
is_matching_prefix <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}

# Removes last element of vector with length x
remove_last_element <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}

remove_last_element(c("statistics", "data science", "data", "199"))

# rep overwrites each element in vector x with y element length(x) times
overwrite <- function(x, y) {
  rep(y, length.out = length(x))
}

overwrite(1:100, c(1,2))

# Checks if number of chars. of x is even (TRUE) or odd (FALSE)
is_even <- function(x) {
  nchar(x) %% 2 == 0
}

is_even("statistic")

```

4. Create `both_na()`, a function that takes two vectors of the same length and 
returns the number of positions that have an NA in both vectors.  Hint: first 
create two vectors like `x <- c(1, 2, 3, NA, NA)` and `y <- c(NA, 1, 2, 3, NA)` 
and write code that works for `x` and `y`, then turn it into a function that can 
handle any `x` and `y`.

```{r test two vectors}
x <- c(1, 2, 3, NA, NA)
y <- c(NA, 1, 2, 3, NA)

# need to check if vectors are same length

both_na <- function(x, y) {
  sum(is.na(x) & is.na(y))
}

is.na(x)
sum(is.na(x))

is.na(y)
sum(is.na(y))

sum(is.na(x) + is.na(y))

both_na(x, y)
```


5. Use a `map` function to find the mean of every column in `mtcars`.

```{r mean fxn}
map_dbl(mtcars, mean)
```


6. Use a `map` function to find the number of unique values in each column of
`msleep`.

```{r num unique vals fxn}
map_int(msleep, n_distinct)
```


7. Use a `map` function to find the type of each column in `msleep`.

```{r column type fxn}
map_chr(msleep, typeof)
```


8. Explain what is meant by "clean coding" and include specific references to how 
the lengths of variable and function names relate to scope.

- Long scope: functions short, variables long

- Small scope: functions long, variables short

9. Brainstorm a tedious task (this could be something relevant to your project 
or something else you think of) and create a function to automate this task.

```{r homemade greetings fxn}

# Homemade greetings fxn that gives message based on time of day (for server)
greetings <- function(time = Sys.time()) {
  hour <- parse_number(str_sub(time, 12, 13))
  # hour(time)
  if (hour < 12) {
    print("Good morning!")
  } else if (hour < 17) {
    print("Good afternoon!")
  } else {
    print("Good evening!")
  }
}

Sys.time()
greetings()

greetings(time = "2019-04-29 09:30:00 CDT") # 9:30 AM (Morning time)
greetings(time = "2019-04-29 23:30:00 CDT") # 11:30 PM (Evening time)

```

