---
title: "Lab 01 - Hello R!"
subtitle: "due 8-27 at 11:59 PM"
author: "Scott Burstein"
date: "08/21/2020"
output:
  html_document: default
  pdf_document: default
---

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(datasauRus)


datasaurus_dozen %>%
  count(dataset) %>%
  print(13)
```

# Exercise 1

The Datasaurus Dozen dataset has 1846 rows and 3 columns. The variables included in the data frame are dataset, x and y.

# Exercise 2

plot of data in the dino dataset :

```{r plot-dino, fig.height=5, fig.width=5}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")
ggplot(data = dino_data, mapping = aes(x = x, y = y)) + 
  geom_point()
```

correlation between `x` and `y` in  dataset 

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

# Exercise 3

Plot of y vs x for the star dataset. Also calculated correlation coefficient between x and y for the dataset.

```{r plot-star, fig.height=5, fig.width=5}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")
ggplot(data = star_data, mapping = aes(x = x, y = y)) + 
  geom_point()
```

This dataset creates a plotted star shape graph.X values are distributed from approximately 20 to 90, where y values distributed from approximately 13 to 90.


```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```

The correlation coefficient for the star dataset is -0.063, which is effectively equal ot the correlation coefficient for the dino dataset.

# Exercise 4

```{r plot-circle, fig.height=5, fig.width=5}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")
ggplot(data = circle_data, mapping = aes(x = x, y = y)) + geom_point()
```
This dataset plots a circle shape. The X and Y values have similar distributions to the dino and star visualizations, which makes sense given that all of the datasets have the same summary statistics.

```{r cor-circle}
circle_data %>%
  summarize(r = cor(x, y))
```

Once again, the correlation coefficient for this dataset is equal to the correlation coefficient's of the previous visualizations. For the circle data, r = -0.0683.

# Exercise 5

```{r plots-all, fig.height=10, fig.width=6}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) + geom_point() + 
  facet_wrap(~ dataset, ncol = 3) +
  theme(legend.position = "none")
```
In the table of charts, all 13 of the visualizations have the same distributions and summary statistics, however they all portray different images (away, bullseye, circle, dino, dots, h_lines, high_lines, slant_down, slant_up, star, v_lines, wide_lines and x_shape).

```{r cors-all}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y))
```
Furthermore, all of these datasets have basically the same r value correlation coefficient of ~ -0.065 ± 0.003.
