---
title: "HW 02 - Statistical inference"
author: "Scott Burstein"
date: ''
output:
  word_document: default
  html_document: default
  pdf_document: default
subtitle: due Tue, Oct 20 at 11:59p
---

# Configure Git

```{r eval = FALSE}
library(usethis)
use_git_config(user.name= "sburstein", user.email="scott.burstein@duke.edu")
```

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
``` 

```{r read-in-data, message = FALSE}
licorice <- read_csv("data/licorice.csv")
colnames(licorice)
```

### Exercise 1

```{r ex1-rm-NA}
mstp_30 <- licorice %>%
  select(pacu30min_throatPain) %>%
  filter(pacu30min_throatPain != "NA")
```

```{r ex1-null-hyp}
mean(mstp_30$pacu30min_throatPain)
```

$H_o: \mu = 0.648$
Null hypothesis is that the mean `pacu30min_throatPain` in all patients
is equal to 0.648.

$H_a: \mu \neq 0.648$
Alternative hypothesis is that the mean `pacu30min_throatPain` in all patients
is not equal to 0.648.

```{r sim-based-CI}
set.seed(123)
n_sims <- 5000
stat <- numeric(n_sims)
for (i in 1:n_sims) {
  indices <- sample(1:nrow(mstp_30), replace = T)
  boot_mean <- mstp_30 %>%
    slice(indices) %>%
    summarize(boot_mean = mean(pacu30min_throatPain)) %>%
    pull()
  stat[i] <- boot_mean
}
boot_means <- tibble(stat = stat)
```

```{r ex-1-interval}
bootstrap_interval <- boot_means %>%
  summarize(lower = quantile(stat, 0.025),
            upper = quantile(stat, 0.975))
bootstrap_interval
```

The 95% confidence interval for the mean height in meters is (0.494, 0.811) 
according to bootstrap method

From simulated based approach can be 95% confident that the true average/mean 
throat pain of patients 30 minutes after arrival in PACU is between the bounds 
of 0.494 and 0.811 out of 10 on a numeric scale (0-10) 
(Based on mean probabilities).

```{r CLT-confidence-interval}
t.test(licorice$pacu30min_throatPain, na.rm=TRUE, conf.level = 0.95)
```
The 95% confidence interval for the mean height in meters is (0.487, 0.809) 
according to CLT.

From CLT can be 95% confident that the true average/mean throat pain of patients 
30 minutes after arrival in PACU is between the bounds of 0.487 and 0.809 out 
of 10 on a numeric scale (0-10) (Based on mean probabilities).


These two methods yield very similar bounds, meaning that using the central 
limit theorem and the simulation method, we can be confident in both instances 
that we do not reject the null hypothesis that the mean `pacu30min_throatPain` 
in all patients is equal to 0.648, since this value lies in both 95% confidence 
intervals constructed.

### Exercise 2

$H_o: \mu = 26$
Null hypothesis is that the BMI among sample patients is not statistically 
different in a significant way from the mean BMI among all German adults of 
26 kg/m^2.

$H_a: \mu \neq 26$
Alternative hypothesis is that the BMI among sample patients statistically 
differs from the mean BMI among all German adults of 26 kg/m^2 in a 
significant manner.

```{r filter-BMI}
pre_BMI <- licorice %>%
  select(preOp_calcBMI)
```


```{r ex-2_simulated-null-dist}
set.seed(123)
n_sims <- 5000
stat <- numeric(n_sims)
for (i in 1:n_sims) {
  indices <- sample(1:nrow(pre_BMI), replace = T)
  boot_mean <- pre_BMI %>%
    slice(indices) %>%
    summarize(boot_mean = mean(preOp_calcBMI)) %>%
    pull()
  stat[i] <- boot_mean
}
boot_means <- tibble(stat = stat)
```

```{r ex-2-shift}
boot_mean <- boot_means %>%
  summarize(mean = mean(stat)) %>% 
  pull()
boot_means <- boot_means %>% 
  mutate(shifted_means = stat + (26 - boot_mean))
```


```{r ex-2 visualization}
ggplot(data = boot_means, aes(x = shifted_means)) +
  geom_histogram(
    binwidth = 0.05, 
    color = "darkblue", 
    fill = "skyblue") +
  labs(
    x = "Mean BMI", 
    y = "Count") +
  geom_vline(xintercept = mean(boot_means$stat), 
             lwd = 1, color = "red")
```


```{r ex-2-interval}
bootstrap_interval <- boot_means %>%
  summarize(lower = quantile(stat, 0.025),
            upper = quantile(stat, 0.975))
bootstrap_interval
```

No, there is not evidence that the mean BMI among such patients differs from the 
mean BMI among all German adults of 26 kg/m^2, since 26 falls within the 95% 
confidence interval of (25.04, 26.15) which the bootstrap simulation shows. 
Thus, we fail to reject the null hypothesis that the BMI among sample patients 
is not statistically different in a significant way from the mean BMI among all 
German adults of 26 kg/m^2.

### Exercise 3

$H_o:$
Null hypothesis is that there is not a relationship between ASA classification 
and having any throat pain after surgery among all patients undergoing surgeries 
that require intubation.

$H_a:$
Alternative hypothesis is that there is some relationship between ASA 
classification and having any throat pain after surgery among all patients 
undergoing surgeries that require intubation.

```{r ex3-rm-NA}
asa_rm_na <- licorice %>%
  select(preOp_asa, pacu30min_throatPain) %>%
  filter(pacu30min_throatPain != "NA")
```

```{r ex-3-table}
ASA_any_pain <- asa_rm_na %>%
  mutate(pain_present = ifelse(pacu30min_throatPain > 0, 1, 0))  %>%
  select(preOp_asa, pain_present)
ASA_any_pain
```
```{r pain-by-asa-summary}
mean_pain_by_asa <- ASA_any_pain %>%
  group_by(preOp_asa) %>%
  summarise(mean_pain = sum(pain_present)/n()) %>%
  ungroup()
mean_pain_by_asa
```

```{r chisq-test}
chisq.test(table(ASA_any_pain$preOp_asa, ASA_any_pain$pain_present))
```

The Pearson's Chi-squared test shows that p-value = 0.2985 for the α level of 
0.05, thus we fail to reject the null hypothesis that there is not a relationship 
between ASA classification and having any throat pain after surgery among all 
patients undergoing surgeries that require intubation.

### Exercise 4

$H_o:$
Null hypothesis is that there is not a relationship between ASA classification 
and having any throat pain after surgery among all patients undergoing surgeries 
that require intubation.

$H_a:$
Alternative hypothesis is that there is some relationship between ASA 
classification and having any throat pain after surgery among all patients 
undergoing surgeries that require intubation.

```{r ex-4}
t.test(licorice$pacu30min_throatPain ~ licorice$treat, data = licorice,
mu = 0,
var.equal = FALSE, alternative = "greater", conf.level = 0.95)
```


From the Welch Two Sample t-test performed, it is evident that the mean of 
`pacu30min_throatPain` for patients who received licorice treatment is 
significantly different from (less than) the mean of `pacu30min_throatPain` for 
patients who received placebo treatment.

mean of `pacu30min_throatPain` for patients who received licorice treatment 
= 
0.274

while mean of `pacu30min_throatPain` for patients who received placebo treatment 
= 
1.026

The p-value = 1.804e-06, which is less than α level of 0.05.

Thus, we have sufficient evidence to reject the null hypothesis (that there 
is not a relationship between ASA classification and having any throat pain 
after surgery among all patients undergoing surgeries that require intubation).


### Exercise 5

```{r ex5-rm-NA}
treat_rm_na <- licorice %>%
  select(treat, pacu30min_throatPain) %>%
  filter(pacu30min_throatPain != "NA")
```

```{r ex-5-data-wrangle}
treatment_any_pain <- treat_rm_na %>%
  mutate(pain_present = ifelse(pacu30min_throatPain > 0, 1, 0))  %>%
  select(treat, pain_present)
treatment_any_pain
```
```{r pain-by-treats-summary}
mean_pain_by_treat <- treatment_any_pain %>%
  group_by(treat) %>%
  summarise(mean_pain = sum(pain_present)/n()) %>%
  ungroup()
mean_pain_by_treat
```

```{r total-prop-pain}
total_prop_pain <- treatment_any_pain %>%
  summarise(mean_pain = sum(pain_present)/n())
total_prop_pain
```

From the total population, the proportion of patients who experience pain is 
0.275.

Null Hypotheses (Ho): Probability pain for licorice group = Probability pain 
for placebo group

The proportion of licorice treatment patients that had pain 30 minutes after 
surgery is not significantly different from the proportion of placebo sugar 
solution patients that had pain 30 minutes after surgery. 

There is no measurable and thus meaningful deviation from the total proportion 
of patients who reported pain 30 minutes after surgery. 

Alternative Hypotheses (Ha): Probability pain for licorice treatment group is 
less than Probability pain for placebo treatment group.

The proportion of licorice treatment patients that had pain 30 minutes after 
surgery is significantly lower than the proportion of placebo sugar solution 
patients that had pain 30 minutes after surgery. 

There is a measurable and thus meaningful difference in probability of 
patients who reported pain 30 minutes after surgery for these two groups.

```{r bootstrap_estimate}
set.seed(18)
number_sims <- 5000
outcomes <- c("pain", "no pain")
simulated_props <- numeric(number_sims)
for (i in 1:number_sims) {
  sim <- sample(outcomes, 233, prob = c(0.2746781, 1 - 0.2746781 ), replace = TRUE)
  simulated_props[i] <- sum(sim == "pain") / length(sim)
}
```

```{r ex5-viz}
simulated_props_data <- tibble(props = simulated_props)
ggplot(simulated_props_data, 
    aes(x = props)) + 
  geom_histogram(bins = 30) + 
  geom_vline(xintercept = (0.1880342), color = "blue") + #licorice group
  geom_vline(xintercept = (0.2746781), color = "red") + #total population
  geom_vline(xintercept = (0.3620690), color = "green") + #placebo group
  labs(x = "Simulated Proportion assuming Ho",
       title = "Proportion of Patients who Experienced Pain")
```
```{r ex5-p-value-licorice}
simulated_props_data %>%
  filter(props <= (0.1880342)) %>%
  summarize(p_value = n() / nrow(simulated_props_data))
```
p-value for licorice treatment group is significantly smaller than 0.05 which is
sufficient evidence to reject the null hypothesis while using a  95% confidence 
interval.

```{r ex5-p-value-placebo}
simulated_props_data %>%
  filter(props >= (0.3620690)) %>%
  summarize(p_value = n() / nrow(simulated_props_data))
```

Similarly, the p-value for licorice treatment group is significantly smaller 
than 0.05 which is also sufficient evidence to reject the null hypothesis 
when using a 95% confidence interval.

### Exercise 6

Yes, based on my analysis, I think that licorice gargle prior to surgery is 
effective in reducing post-intubation sore throat. 

From the Welch Two Sample t-test in exercise 4, it is evident that the mean of 
`pacu30min_throatPain` for patients who received licorice treatment is 
significantly lower than the mean of `pacu30min_throatPain` for patients who
received placebo treatment.

In addition, exercise 5 proved that the proportion of licorice treatment 
patients that had pain 30 minutes after surgery is significantly lower than the 
proportion of placebo sugar solution patients that had pain 30 minutes after 
surgery. 

From these statistical tests, it is clear that there is a measurable and thus 
meaningful deviation in both prevalence and intensity of pain between the 
licorice treatment patients and the placebo treatment patients 30 minutes after 
surgery.

***
**For Ex. 7-12 Given That:**
The mean BMI among patients receiving licorice solution was 25.6 kg/m^2 and the 
mean BMI among patients receiving sugar solution placebo was 25.6 kg/m^2. 
In assessing whether there is a difference in mean BMI between the two treatment 
groups using a CLT-based approach, the researchers obtained a p-value of 0.925.

### Exercise 7

**Statement:**
If there is truly no difference in mean BMI between these two groups, then the 
probability of seeing a difference in BMI as large as our observed difference 
or even larger is approximately 0.925.

TRUE

### Exercise 8

**Statement:**
Assuming α = 0.05, then our p-value of 0.925 would be strong evidence that there 
is no difference in the mean BMI between the two treatment groups.

TRUE

When the  p-value is greater than the stated alpha level, we fail to reject the 
null hypothesis. Hence, in this instance, where p-value is 0.925 and α = 0.05,
we have sufficient evidence to NOT reject the null hypothesis, that there 
is no difference in the mean BMI between the two treatment groups. This is,
in a logical sense, effectively supporting the null hypothesis.

### Exercise 9

**Statement:**
The probability that we have made a type 2 error is less than 10%.

FAlSE

A type II error occurs when we fail to reject a null hypothesis that is false.
Beta (prob type 2 error = 1-power of test), can not be extrapolated from p-value 
alone (all that is given).

### Exercise 10

**Statement:**
If we were to repeatedly construct 95% confidence intervals for the difference 
in mean BMI in the same way from the original population, then approximately 
95% of those intervals would truly contain the true population difference in 
means.

FALSE

We are 95% confident that the true difference in mean BMIs lies within the 
bounds for each 95% confidence interval constructed for the difference in mean 
BMIs. However, we cannot conclude that 95% of the constructed intervals 
contain the true difference in mean BMIs.

### Exercise 11

**Statement:**
With our given p-value, we know for sure that constructing a 95% confidence 
interval for the mean difference in BMIs would definitely contain 0.

FALSE

We cannot be certain that the 95% confidence interval contains 0.
By definition, we can only be 95% confident that the true difference lies within
the bounds specified by the 95% confidence interval. Furthermore, since the 
p-value is 0.925, it is possible that events as or more extreme as the null 
hypothesis occur. We consequently cannot be certain that the 95% confidence 
interval contains 0.

### Exercise 12

**Statement:**
If we instead found a p-value of 0.021, then at the α = 0.05 level, we would 
have enough evidence to conclude that there is a difference in mean BMI between 
the two treatment groups.

TRUE
