---
title: "AE 14"
subtitle: "Birth Weights"
date: "Due: October 21, 11:59PM"
author: "SCOTT BURSTEIN"
output: 
  html_document
---

# Configure Git

```{r eval = FALSE}
library(usethis)
use_git_config(user.name= "sburstein", user.email="scott.burstein@duke.edu")
```

# Load Packages

```{r packages, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(broom)
```

The Child Health and Development Studies investigated 1,192 pregnancies between 
1960 and 1967 among women on  the  Kaiser  Foundation  Health  Plan  in  San 
Francisco. Researchers  were  interested  in  modeling  the  weight  of the 
infants in ounces (`bwt`) using the pregnancy time in days (`gestation`), 
whether the child was first born (`parity`, 0 means first pregnancy), the 
mother’s height in inches (`height`), and whether the mother was a smoker 
(`smoke`, 1 means smoker). Data is from *OpenIntro Statistics (4th edition)* 
and available in the `births.csv` file.

```{r load-data, message = FALSE}
births <- read_csv("data/births.csv")
```

## Multiple Regression

1. Fit a multiple linear regression model predicting `bwt` with the main effects
`gestation`, `parity`, `height`, and `smoke` and display the R output in tidy
format.

```{r reg1}
m_main <- lm(bwt ~ gestation +parity + height + smoke, data = births)
tidy(m_main) %>% 
  select(term, estimate) %>% 
  mutate(estimate = round(estimate, 5))
```

2. Write out the equation of the estimated regression line.

$$\widehat{weight} = -85.42 + 0.44~gestation -3.46~parity + 1.34~height - 8.47~smoke$$

3. Interpret each of the estimated model coefficients.

- **Intercept**: Babies with a pregnancy length of 0, that are first-born,
born to nonsmoking mothers with a height of 0 are predicted to have a birth weight 
of -85.42 ounces.
- **Gestation**: For every one day increase in pregnancy length, the birth weight 
of a baby is predicted to increase, on average, 0.44 ounces holding all else constant.
- **Smoke**: Babies born to smoking mothers are predicted to have a birth weight that is 
8.47 ounces less than babies born to nonsmoking mothers, holding all else constant.

4. Report $R^2$ and provide an interpretation in the context of the problem. Why
might the adjusted $R^2$ be preferred over $R^2$?

```{r glance}
glance(m_main)
```


5. Construct a 95% confidence interval for $\beta_{height}$ and provide a 
concise, careful interpretation in the context of the problem.

```{r tidy}
tidy(m_main, conf.int = TRUE, conf.level = 0.95) %>% 
  filter(term == "height")

height_baby <- lm(bwt ~ height, data = births)
tidy(height_baby)
```

We are 95% confident that for a one inch increase in the mothers height, the 
birthweight is predicted to increase, on average, between 0.980 and 1.69 ounces,
holding all else constant.

## Multiple Regression with Interaction

6. Fit a multiple linear regression model predicting `bwt` with main effects for
`gestation`, `smoke` and their interaction.

```{r model2}
m_int <- lm(bwt ~ smoke + gestation + gestation * smoke, data = births)
tidy(m_int)
```


7. Write out the model equations for smokers and nonsmokers.


$$\widehat{Weight} = 20.3 - 73.5~smoke + 0.367~gestation + 0.234~smoke \times gestation$$

**Nonsmokers:**
$$\widehat{Weight} = 20.3 + 0.367~gestation$$

**Smokers:**
$$\widehat{Weight} = -53.21 + 0.603 \times gestation$$
8. Visualize the linear model on a single set of axes.

```{r viz-model}
ggplot(data = births,
       mapping = aes(x = gestation, y = bwt, color = factor(smoke))) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE, fullrange = TRUE)
```

