---
title: "In-class Exercise 10-27"
subtitle: "Medical School Admission"
author: "SCOTT BURSTEIN"
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r load-packages, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(broom)
```

We will work with a dataset containing information on 55 medical school
applications from undergraduate students. We are interested in predicting
`Acceptance` based on `Sex` and `MCAT` score.

```{r load-data-and-packages, message = FALSE}
med_school <- read_csv("data/medschool.csv")
```

## Logistic Regression

1. Perform a brief EDA of the relationship between sex and acceptance and MCAT
score and acceptance.

```{r ex-1}

ggplot(med_school,
       aes(x = Sex, fill = factor(Acceptance))) + 
  geom_bar(position = "fill")

ggplot(med_school,
       aes(x = MCAT, y = Acceptance, color = Sex)) + 
  geom_jitter(height = 0.01)

```

 
2. Fit a logistic regression model with `Acceptance` as the response and `MCAT`
and `Sex` as predictors. Report the model output in tidy format.

```{r ex-2}
m_mcat <- glm(Acceptance ~ MCAT + Sex, data = med_school, family = "binomial")
tidy(m_mcat)
```


3. Write out the estimated model.

$$ \text{log}\left(\frac{\hat{p}}{1 - \hat{p}}\right) = -8.95 + 0.267~MCAT - 1.07~Sex$$

$$ \hat{p} = \frac{e^{-8.95 + 0.267~MCAT - 1.07~Sex}}{1 + e^{-8.95 + 0.267~MCAT - 1.07~Sex}}$$

4. Write out the model for men.

$$ \text{log}\left(\frac{\hat{p}}{1 - \hat{p}}\right) = -8.95 + 0.267~MCAT - 1.07$$

5. Write out the model for women.

$$ \text{log}\left(\frac{\hat{p}}{1 - \hat{p}}\right) = -8.95 + 0.267~MCAT$$

6. What is the interpretation of $\hat{\beta}_{mcat}$?

Holding all else constant, the odds ratio of admission for an additional point 
on the MCAT is $e^{0.267} = 1.306$.

Holding all else constant, for every one unit increase in MCAT core, the odds of admission is multiplied by a factor of 1.306.

7. What is the interpretation of $\hat{\beta}_{sex}$?

Holding all else constant, the odds ratio of acceptance to medical school for 
males versus females is $e^{-1.0622} = 0.3443$

8. What is your prediction for the log odds of acceptance for a male student
with an MCAT of 30?

```{r ex-8}
newdata <- tibble(Sex = "M", MCAT = 30)
pred_log_odds <- augment(m_mcat, newdata = newdata) %>% 
  pull(.fitted)
pred_log_odds
```

9. What is your prediction for the acceptance probability for a male student
with an MCAT of 30?

```{r ex-9}
pred_prob <- exp(pred_log_odds) / (1 + exp(pred_log_odds))
pred_prob
```
