---
title: "In-Class Notes 10-1"
subtitle: "Central Limit Theorem"
author: "Scott Burstein"
date: "10-1"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Get started

### Clone a repo + start a new project 

Go to the `inclass03-[GITHUB USERNAME]` repo, clone it, and start a new project in RStudio. See earlier labs and application exercises for more detailed instructions.

### Configure git

Run the following code to configure Git. Fill in your GitHub username and the email address associated with your GitHub account. 

```{r eval = FALSE}
library(usethis)
use_git_config(user.name = "sburstein", user.email ="scott.burstein@duke.edu")
```

We will be using the `pokemon` dataset, which contains information about 45
randomly selected Pokemon (from all generations). You may load in the dataset
with the following code (be sure to set `eval` to be `TRUE` in the following R
chunk).

```{r load-data, message = FALSE, eval = TRUE}
library(tidyverse)
pokemon <- read_csv("data/pokemon.csv")
```

### Exercise 1

Construct and interpret a 95% confidence interval for the mean height in meters 
(`height_m`) of Pokemon species by using the Central Limit Theorem.

```{r confidence-interval}
t.test(pokemon$height_m, conf.level = 0.95)
```
The 95% confidence interval for the mean height in meters is (0.798, 1.891).

Are 95% confident that the true average/mean height of Pokemon species is 
between the bounds of 0.798 and 1.891 meters. (Based on mean probabilities)

```{r ex1-by-hand}
summary_stats <- pokemon %>%
  summarize(xbar = mean(height_m),
            s = sd(height_m),
            n = n())

tstar <- qt(0.975, df = summary_stats$n - 1)

se <- summary_stats$s / sqrt(summary_stats$n)

summary_stats$xbar + c(-1, 1) * tstar * se
```


### Exercise 2

The average height of a human is 1.65 meters. Determine if there is enough 
evidence to suggest that the true mean height of Pokemon is **different** from 
1.65 meters using the Central Limit Theorem. State your null and alternative 
hypotheses, the distribution of the statistic under the null hypothesis, the 
p-value, statistical decision, and a conclusion in the context of the problem.

$H_o: \mu = 1.65$

$H_a: \mu \neq 1.65$

```{r ex-2-use-function}
t.test(pokemon$height_m, mu = 1.65)
```

Fail to reject the null hypothesis. Since the p-value is large, it is not 
surprising to see values as or more extreme as 1.65 assuming null hypothesis is 
true.

```{r ex2-by-hand}
summary_stats <- pokemon %>%
  summarize(xbar = mean(height_m),
            s = sd(height_m),
            n = n())
summary_stats

se <- summary_stats$s / sqrt(summary_stats$n)
se

t <- (summary_stats$xbar - 1.65) / se     #mu = 1.65
t

2 * pt(t, df = summary_stats$n - 1) #2 * left tail value bc both sides

```

Testing philosophy: Assume that the null hypothesis is true, estimate how 
surprising it is to see values as or more extreme as observed sample data.

ASSUMPTIONS TO MAKE: Null Hypothesis is TRUE

### Exercise 3

Now determine if there is enough evidence to suggest that the true mean height 
of Pokemon is **lower** than 1.65 meters using the Central Limit Theorem. 
State your null and alternative hypotheses, the distribution of the statistic 
under the null hypothesis, the p-value, statistical decision, and a conclusion 
in the context of the problem.

Since we failed to reject the null hypothesis, and there is enough evidence to 
suggest that the true mean height of Pokemon is lower than 1.65 meters using 
the Central Limit Theorem, we came to the wrong result, and made a type II error 
(failing to reject a FALSE null hypothesis)

```{r ex3}
##Answers and values above
```


### Exercise 4

Suppose that the true mean height of Pokemon is 1.4 meters. In your conclusions 
above did you make the correct decision, a Type I error or a Type II error?
Explain.

```{r ex4}
##Answers and values above
```

